API design for machine learning software: experiences from the scikit-learn project

### 2013-06-25 paper.tex
comparison to VW, SofiaML, Weka (CLI tools)
Comparison section needs heavy editing.

\textcolor{red}{TODO: compare with weka, R, vowpal rabbit, ... ?}\\
\textcolor{red}{TODO: emphasize scikit-learn's place in the overall scipy ecosystem, and its role in spurring the development of more basic tools.}\\
\textcolor{red}{TODO: some packages start to adopt our own API.}

@@

@@

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim}.
Although its stated goal is not machine learning,
it exemplifies API design for scalable processing of ``big data''
which takes a different approach from the comparable models in \sklearn.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorization part of \sklearn
also uses iterables to some extent,
the emphasis is still on (mini-)batch processing,
even in the stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and makes effective use of the abundance of memory in modern computers.
We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.

>>>>

\textcolor{red}{TODO: emphasize scikit-learn's place in the overall scipy ecosystem, and its role in spurring the development of more basic tools.}\\

@@

Many machine learning packages exist, with new ones still often appearing,
that are intended to be used as command-line tools
(and sometimes do not offer any type of API).
Examples include Weka \citep{hall2009weka}, SofiaML \citep{sculley2010web}
and Vowpal~Wabbit \citep{langford2009sparse}.
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need to do programming,
and will do so in a variety of languages.
We argue that this makes it harder for researchers and developers
to exchange software and collaborate on software.

@@

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim},
which exemplifies API design for scalable processing of ``big data''.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorization part of \sklearn
also uses iterables to some extent,
the emphasis is still on (mini-)batch processing,
even in the stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and makes effective use of the abundance of memory in modern computers.
We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.
